---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r Load Data}
myDataf <- read.csv(unz("./activity.zip", "activity.csv"), header = TRUE,
                    colClasses = c("integer", "Date", "numeric"), sep = ",",
                    na.strings = "NA", quote = "\"")
```

## What is mean total number of steps taken per day?

Histogram of the total number of steps taken each day

```{r Histogram steps x day}
myDataf.steps <- aggregate(steps ~ date, myDataf, sum)
with(myDataf.steps, barplot(myDataf.steps$steps,
                            names.arg=myDataf.steps$date, xlab = "Date", ylab ="Steps",
                            col="red", density = 20))
```

Mean and median number of steps taken each day
```{r Mean + Median steps x day}
mean(myDataf.steps$steps, na.rm = TRUE)
median(myDataf.steps$steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

Time series plot of the average number of steps taken
```{r  Plot Avg steps}
steps.interval <- aggregate(steps ~ interval, activity, mean)

```


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
